---
title: "Chapter 2 Lab"
author: "Tianwei Liu"
output: pdf_document
---

## Preparation

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
## If any of the packages do not exist in your environment (meaning you have not previously downloaded
## them) you should first install them using (for example):
## install.packages("haven") 

## Load packages used in this session of R
require(haven) 
require(knitr)
require(car)

opts_chunk$set(echo = TRUE)
options(digits = 3)

#insert path to your working directory below
opts_knit$set(root.dir ="~/Desktop/GU/Stats/Lab1")   
# You cannot use setwd() with knitr, so use this command to set root directory where data is saved
dta <- read_dta("Ch2_lab_survey_data.dta")
```

#### 1) Use the following to create dummy variables for Arlington and Prince William Counties.  How many observations are from each county?

```{r tidy = FALSE}
dta$Arlington <- (dta$precinct == "AR49" | dta$precinct == "AR22" | dta$precinct == "AR2" | 
  dta$precinct == "AR18" | dta$precinct == "41" | dta$precinct == "4" | dta$precinct == "16" | 
  dta$precinct == "17" | (dta$precinct == "2" & dta$state == 4 & !is.na(dta$state)) | 
  dta$precinct == "31" | dta$precinct == "48")

dta$PrinceWilliam <- (
  dta$precinct == "PW 101" | dta$precinct == "PW 104" | dta$precinct == "PW101" | 
  dta$precinct == "PW 401" | dta$precinct == "PW104" | dta$precinct == "PW402" | 
  dta$precinct == "PW406"|    dta$precinct == "401" | dta$precinct == "402" | 
  (dta$precinct == "104" & dta$state == "4" & !is.na(dta$state))
)
## How many observations are from Arlington
table(dta$Arlington)
## How many observations are from Prince William
table(dta$PrinceWilliam)

```

Therefore, there are 475 observations in Arlington and 188 observations in Prince William County. 

#### 2) Create dummy variables for each state/DC.  How many observations are in DC, Maryland, Ohio and Virginia?

```{r tidy = FALSE}
dta$DC <- (dta$state == 1)
table(dta$DC)
dta$MD <- (dta$state == 2)
table(dta$MD)
dta$OH <- (dta$state == 3)
table(dta$OH)
dta$VA <- (dta$state == 4)
table(dta$VA)
```

Overall, there are 768 observations in DC; 369 observations in MD; 547 observations in OH; and 664 observations in VA.

#### 3)	Convert the year_born variable into age.  Be sure to check for and correct for data errors.  What is the average age of all observations in the data set?  The minimum and maximum?

```{r}
dta$age <- 2016 - dta$year_born
summary(dta$age)
dta$age[dta$year_born <= 1920] <- NA
dta$age[dta$year_born > 2016] <- NA
summary(dta$age)
```

After correction, the average age is 41. Minimum is 17 and maximum is 95.

#### 4)	What is the distribution of the gender variable?  Create a male dummy variable and indicate the distribution of this variable. Compare distribution of your male variable to the distribution of the gender variable.

```{r}
dta$male <- (dta$gender == 1)
table(dta$male)
table(dta$gender)
```

Other than male, there are 5 people selected "other".

#### 5)	Provide descriptive stats for Trump and Clinton feeling thermometer.  Is there anything you need to adjust?
```{r}
summary(dta$therm_trump)
summary(dta$therm_clinton)
dta$therm_clinton[dta$therm_clinton > 100] <- NA
summary(dta$therm_clinton)
```

#### 6)	What is the distribution of the education variable?  Is there any adjustment you would need to make if you will use this as a continuous variable in a regression model?
```{r tidy = FALSE}
table(dta$education)
dta$education[dta$education == 4] <- NA
dta$education[dta$education == 5] <- 4
dta$education[dta$education == 6] <- 5
dta$education[dta$education == 7] <- 6
summary(dta$education)
```

#### Additional Comments: It may be convenient to save the data, as we'll use this in the lab for next chapter. Here, we save the data as .Rdata format.

```{r tidy = FALSE}
# save the data as .Rdata format
save.image(file = "data_chapter3_lab.Rdata")
# To load this .Rdata file late,  use the load function: load("data_chapter3_lab.Rdata")
```
